---
title: "DrexelRSeminarSummer19"
output: html_document
---
Ok, this is my first effort at making a course in RStudio.

To get to this point you've done the basics of getting R installed, RStudio, and connecting it to Git.

You will want to Fork this Repo to make your own. 

To get started you'll need to install some packages, they are:

here, pacman, tidyverse

In order to get started, you have an assignment: 
I have provided in the data folder several files that include the test responses and the grades for two classes, as well as a grading key. 

The first part of this course will get you going and require you to clean up this data - NO EXCEL!!!!

I want a script that imports the data, combines it all into one dataframe, and that correlates the test score with the quiz 1 grade and quiz 2 grade.  


```{r SetupEnvironment, echo=FALSE}
pacman::p_load(here, tidyverse, lubridate, CTT)


TestDf = read_csv("data/AnonTestResponses - TestData1709to1812.csv")
Grades17 = read_csv("data/AnonCourseGrades1712.csv")
Grades18 = read_csv("data/AnonCourseGrades1812.csv")
TestKey =  read_csv("data/TestKey.csv", col_names = F)



```

Here I'll tidy up the TestDf.  Renaming the variables, and getting the time as a time and the section as a factor.  Then I don't really care about all of the text, I want the first letter in each of the TestQ columns.

```{r TidyTestDF, echo= FALSE}

TestDf %>%
  rename( ID = `What is your 8 digit student id?`,
          Section = `Which lecture section are you in?`) -> TestDf


#There doesn't seem to be a good way to do this in dplyr...
TstNames = data.frame(Tst = "TestQ", Item = seq(1:30))
TstNames %>% unite(ItemNum, Tst, Item) -> TstNames
colnames(TestDf)[4:33] <- t(TstNames)

TestDf %>%
  mutate(Timestamp = mdy_hms(Timestamp),
         TestDate = date(Timestamp),
         Section = factor(Section)) -> TestDf
  
TestDf %>%
  mutate_at(vars(starts_with("TestQ")), list(~str_extract(., "[A-E]"))) ->TestDf



```

Now, I will tidy up the grades...

```{r TidyGradeDF,echo=FALSE}

Grades17 %>%
  rename(Total = `Total [Total Pts: up to 500] |1403917`,
         Percent = `Running Grade [Total Pts: up to 100] |1447324`,
         HW = `HW [Total Pts: 100] |1447322`, 
         Q1 = `Q1 [Total Pts: 100] |1447316`, 
         Q2 = `Q2 [Total Pts: 100] |1447318`,
         Labs = `Labs [Total Pts: 100] |1447319`) %>%
  select(-`Final [Total Pts: 100] |1447315`) %>%
  mutate(Percent = parse_number(Percent))-> Grades17

Grades18 %>%
  rename(Total = `Total [Total Pts: up to 490 Score] |1704781`,
         Percent = `WeightedTotal [Total Pts: up to 98.5 Percentage] |1704782`,
         HW = `Homework [Total Pts: 100 Score] |1719593`, 
         Q1 = `Quiz1 [Total Pts: 90 Score] |1719589`, 
         Q2 = `Quiz2 [Total Pts: 100 Score] |1719590`,
         Labs = `LabGrade [Total Pts: 100 Score] |1719591`) %>%
  select(ID, Total, Percent, HW, Q1, Q2, Labs) -> Grades18

Grades = bind_rows(Grades17, Grades18)
```

Now that I have them all bound up, I can join the grades and the TestDf

```{r JoinDataFrames, echo= FALSE}
FullData = left_join(Grades, TestDf, by = "ID")

KeyVector = as.vector(TestKey$X1)

FullData %>%
  select(starts_with("TestQ")) %>%
  mutate(Scr = score(., KeyVector)$score) %>%
  select(Scr) -> TestScore
            
FullData = bind_cols(FullData, TestScore)
```


Here I am going to set the pre and the post 

```{r CleanDates, echo= FALSE}
#take a look at the dates that are present
table(FullData$TestDate)

PreDates = c("2017-09-27", "2018-09-24", "2018-09-25", "2018-09-26", "2018-09-27", "2018-10-01")

PostDates = c("2017-12-04", "2018-12-03", "2018-12-05")

FullData %>%
  mutate(PrePost = case_when(
             as.character(TestDate) %in% PreDates  ~ "Pre",
             as.character(TestDate) %in% PostDates ~ "Post")) -> FullData

```

```{r RunCorrelations, echo= FALSE}

FullData %>%
  group_by(PrePost) %>%
  summarise(N = n(), 
            AveTestSc = mean(Scr, na.rm = T), 
            corQ1 = cor(Scr, Q1, use = "pairwise.complete.obs"), 
            corQ2 = cor(Scr, Q2, use = "pairwise.complete.obs"))

```


