---
title: "Chapter 5: Descriptive Statistics"
author: "Virginia Price"
date: "7/1/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lsr)

```


## 5.1 Measures of Central Tendency
"Central tendency" -> measure of the "middle" or "average" of a data set

Here's our data set for this chapter; it includes the teams that played in the playoffs, as well as the margin by which they won.
```{r data-load}
load("../Navarro-data/aflsmall.Rdata")
```
### Mean
The average! Or the "center of gravity" of the data set
notation: 
* total number of observations: N
* each observation = X; individual observations, $X_1, X_2 ... X_N$
* Mean is $\bar{X}$, given by 
    \[\bar{X} = \frac{1}{N}\displaystyle\sum_{i=1}^N{X_i}\]
    
So for 5 observations,
```{r mean-5}
(56+31+56+8+32)/5       #ineffieceint and non-generalizable
sum(afl.margins[1:5])/5 # way better
mean(afl.margins[1:5])  # the best
```

Buuuut R has an inbuilt function for this so let's just find the mean of errything
```{r mean}
mean(afl.margins)
```

### Median
The observation in the middle ("the middle name")

For our 5-observation subset,

8,31,**32**,56,56 -> 32 is the median

Adding a 6th game,
8,14,**31**,**32**,56,56 -> 31.5 is the median (average of the two middle numbers)

R has an inbuilt function for this too:
```{r median}
median( afl.margins )
```

When do we use these different central measures?

* *interval data* -- choose between median and mean! Note that means can be pretty sensitive to outliers. Median is good if you want "typical", mean is good if you want overall
* *ordinal data* -- ranked data, like a Likert scale -- use the median!
* *nominal data* -- that is, data that is not weighted by its number (e.g., words!) - Don't use central measures! Focus on grouping analyses

*Example:* an Australian housing market bank used the mean instead of the median and ended up with a much lower housing:income ratio because it averaged rich peoples' incomes compared it to the median house price. This skewed the dataset from 9:1 -> 5:1!

### Trimmed mean

Dataset: -100,2,3,4,5,6,7,8,9,1

-100 is probably an outlier!
But what if the 'outlier' is -15 instead? Should we still include it? At what point does that data become an outlier?

Use the median, or the "trimmed mean" by discarding the most extreme examples *on both ends*. Generally uses more info than the median to get a better idea of the dataset.

```{r trimmed-mean}
dataset <- c( -15,2,3,4,5,6,7,8,9,12)
mean(dataset)
median(dataset)
mean(dataset, trim=0.1) #trims 10% of the dataset: compare to median!

#For the margins data:
mean(afl.margins,trim=0.05)

```

### Mode
The value that occurs most frequently! R can make frequency tables!

```{r mode-ft}
table( afl.finalists )
```

R *cannot* calculate the mode, but there's functions in the lsr library for that:
```{r mode-lsr}
modeOf(afl.finalists) # [1] "Geelong"
# Finds the thing that appears most frequently
# Geelong has played the most in the finals between 1987-2010

maxFreq( afl.finalists ) # [1] 39
# Tells you the number of times the most popular team appeared

```

## 5.2 Measures of Variability

