---
title: "ANOVA"
author: "Virginia Price"
date: "7/29/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(here, tidyverse, lubridate, lsr)
```

The challenge for next week is to perform an ANOVA on the Scr variable, looking to see if there are mean differences on the basis of the Breakfast grouping variable on the Post data. 

Bonus points to carry out the analysis using a linear model.

Bonus, bonus points if you do an appropriate followup test.

Bonus, bonus, bonus points if you calculate appropriate effect sizes for any differences. 

```{r LoadData}
TestData <-readRDS(file = "CleanTestData2.Rda")
Breakfast <- readRDS(file = "CleanTestData.Rda")
TestData$Breakfast <- Breakfast$Breakfast
rm(Breakfast)
# Read in the datas

# Do the same restructuring as we did for the t-test
PreData <- TestData %>%
  select(ID,PrePost,Scr,Breakfast) %>%
  filter(PrePost == "Pre") %>%
  rename(Scr.Pre = "Scr",
         Breakfast.Pre = "Breakfast")


PostData <- TestData %>%
  select(ID,PrePost,Scr,Breakfast) %>%
  filter(PrePost == "Post") %>%
  rename(Scr.Post = "Scr",
         Breakfast.Post = "Breakfast")

PrePostDf <- left_join(PreData,PostData,by="ID")

```

Data is all set up and ready for ANOVA-ing!

```{r anova-01}
Nova <- aov(
  formula = Scr.Post ~ Breakfast.Post,
  data = PostData
)

summary(Nova)

```

"We cannot reject the null hypothesis that there is a difference in means of the post-test scores based on the grouping variable Breakfast" p>0.885

p-value: how much of the time are we willing to reject the null hypothesis (be wrong)

### Effect size
$ \eta^2 = \frac{SS_{group}}{SS_{residuals}} = \frac{146}{91415}

```{r eta-squared}
etaSquared(x = Nova)
```
0 means no correlation, 1 is perfectly correlated.

"The world is round p<.05"

```{r pairwise-t-test}

```

Bad stuff happens if more than 3 groups. More groups need a more restrictive p-value
If you run more than one pair you run a t-test on, you will probably get a type 1 error good p value. Need to be more constrained. 

Bonferooni says if we run 20 t-tests, take p and divide by 20