---
title: "DrexelRSeminarSummer19Week4"
author: "Eric Brewe"
date: "8/5/2019"
output: html_document
---

```{r setup, include=FALSE}
pacman::p_load(here, tidyverse, lubridate, lsr, broom)
FullData = readRDS(file = "CleanTestData-Fixed.Rda")
FullData %>%
  filter(PrePost == "Post") -> PostData
```

Ok, this week the challenge is to do a linear regression model that predicts the outcome variable (Scr) from a predictor (Quiz1)...using only the post data, this isn't alltogether that interesting.  

```{r LinearModel, echo=FALSE}
Mod1 = lm(Scr ~ Quiz1, data = PostData)
summary(Mod1)
```
The real challenge is understanding what this means....

We get an intercept of 2.2668 (this means if you got a 0 on Quiz 1, your expected Scr would be 2.267)

The coefficient for Q1 is 0.7299 (which incidentally is measured in %). So for every additional % on Q1, we would expect that the Scr goes up by 0.7299 points..Imagine I got a 5% on Quiz 1, you would then expect that my Scr would be 
Scr = 0.7299*(5) + 2.2668 = 5.8813

It also says that the predictor (Quiz1) is a significant predictor and the t-value is 6.397, this is a one sample t-test comparing the value of the coefficient to zero.  t = 0.7299/0.1141 = 6.39  So the t value compares the value of the coefficient to the standard error.   

Chapter 15 does a really nice job of going through the analysis and checking a bunch of assumptions - and you should do all these things prior to publication.  One thing it doesn't do is a nice job of plotting.  So time for a brief dip into ggplot2. I really recommend printing out this cheatsheet.

https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf

run each of the pairs of commands together...this is how ggplot works, you can continue to add layers

```{r Plotting, echo=FALSE}
#This makes tells ggplot what data, and what variables go where...but doesn't plot anything yet.
#aes: "aesthetic mapping"
P1 = ggplot(data = PostData, aes(x = Quiz1, y = Scr))
P1
#This says, take the thing from before and plot it as a scatterplot
P1 = P1 + geom_point()
P1
#This adds the regression line (with the standard error)
P1 = P1 + geom_smooth(method = "lm")
P1
#This cleans up the look.
P1 = P1 + theme_bw()
P1
#This adds a title
P1 = P1 + ggtitle("Eric's Kickass Plot")
P1
#This is how you save a plot, don't run this..
#ggsave("EricKickassPlot.pdf", plot = P1)
```
The last thing you might want to do is to use the model you have on new data. For example, you might have a class the next quarter, and you get a set of scores on quiz 1....how do you expect they will do on the Scr at the end? 

I'll generate a new data set and we can use it to predict. 

```{r Predict, echo=FALSE}
#This generates the new data set...notice the variable has to be the same, eg, Qui1 has to be the name of the variable in order to make the prediction
set.seed(123)
NewQuiz1 = data.frame(Quiz1 = rnorm(30, mean = 65, sd = 15))
ScrPred = predict(Mod1, NewQuiz1)
hist(ScrPred)
mean(ScrPred)
```
So with a quiz 1 score of 65, you might predict a Scr of 49.2 on the post, this does open up other means of evaluating your model (which is what data scientists do a ton of.)

Previously I didn't have the seed set, and one time I ran it and it got a predicted score of 109...this can happen.  So set your seed folks.